Enabling conditions for effective community based marine reserves in small scale fisheries

### 2018-03-20 docs/paper/Villasenor-Derbez.Rmd
Added manuscript to its own folder

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(knitr.table.format = "latex")
```

@@

```{r loadp_kgs}
suppressPackageStartupMessages({
  library(sp)
  library(sf)
  library(ggsn)
  library(tmap)
  library(magrittr)
  library(kableExtra)
  library(broom)
  library(cowplot)
  library(tidyverse)
})

# Load indicator functions
source(file = "../Scripts/abundance.R")
source(file = "../Scripts/richness.R")
source(file = "../Scripts/shannon.R")
source(file = "../Scripts/biomass.R")
source(file = "../Scripts/trophic.R")

# Load analysis functions
source(file = "../Scripts/model.R") #fits the model
source(file = "../Scripts/robust_se.R") #calculates heteroskedastic-robust Standard Errors
source(file = "../Scripts/plot_model.R") #plots effect sizes
```

@@

# Load high-res mexican coastline
coastline_mx <- readRDS("../Data/Spatial/coastline_mx.rds")

@@

# Load ZRP shapefile
zrp <- st_read(dsn = "../Data/Spatial", layer = "ZRP", quiet = T) %>% 
  filter(Name %in% c("El Cabezo", "El Faro")) %>% 
  st_transform(crs = "+proj=longlat +datum=NAD27")

# Load Community reserve shapefile
rescom <- st_read(dsn = "../Data/Spatial", layer = "Res_Com", quiet = T) %>% 
  filter(Nombre == "La_Plana") %>% 
  st_transform(crs = "+proj=longlat +datum=NAD27")

@@

```{r load_data}
fish <- read.csv("../Data/fish_data.csv", strip.white = T, stringsAsFactors = F) %>% 
  filter(Comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  filter(!is.na(RC),
         RC %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
                   "Cabezo - Cabezo (Control)",
                   "Manchon - Manchon (Control)",
                   "El Faro - El Faro (Control)")) %>% 
  filter(!(RC %in% c("El Faro - El Faro (Control)") & Ano == 2012)) %>% 
  mutate(GeneroEspecie = case_when(
    GeneroEspecie == "Cephalopholis cruentatus" ~ "Cephalopholis cruentata",
    GeneroEspecie == "Haemulon macrostomun" ~ "Haemulon macrostomum",
    GeneroEspecie == "Xcochin spp" ~ "Balistes capriscus",
    GeneroEspecie == "Pterois volitans/miles" ~ "Pterois volitans",
    GeneroEspecie == "Mycteroperca spp" ~ "Mycteroperca",
    GeneroEspecie == "Chromis spp" ~ "Chromis",
    GeneroEspecie == "Damisela spp" ~ "Stegastes",
    GeneroEspecie == "Haemulon spp" ~ "Haemulon",
    GeneroEspecie == "Lancero spp" ~ "Acanthurus",
    GeneroEspecie == "Kyphosus sectatrix/incisor" ~ "Kyphosus",
    GeneroEspecie == "Sebastes spp" ~ "Sebastes",
    GeneroEspecie == "Loro spp" ~ "Scarus",
    GeneroEspecie == "Lutjanus spp" ~ "Lutjanus",
    GeneroEspecie == "Mariposa spp" ~ "Chaetodon",
    GeneroEspecie == "Pluma spp" ~ "Calamus pennatula",
    GeneroEspecie == "Sparisoma aurolineatum" ~ "Sparisoma aurofrenatum",
    GeneroEspecie == "Vieja spp" ~ "Bodianus rufus",
    TRUE ~ GeneroEspecie
  ))

invert <- read.csv("../Data/invert_data.csv", strip.white = T, stringsAsFactors = F) %>% 
  filter(Comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  filter(!is.na(RC),
         RC %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
                   "Cabezo - Cabezo (Control)",
                   "Manchon - Manchon (Control)",
                   "El Faro - El Faro (Control)")) %>% 
  filter(!(RC %in% c("El Faro - El Faro (Control)") & Ano == 2012))

load("../Data/conapesca_2000-2015.RData")

cpi <- read.csv("../Data/CPI.csv", stringsAsFactors = F, strip.white = T)

@@

gob <- read.csv("../Data/governance.csv", strip.white = T, stringsAsFactors = T)
```

@@

```{r}
read.csv("../Data/gov_ind.csv") %>% 
  knitr::kable(caption = "\\label{table:gov_ind}Indicators used for the operationalization of the SES framework \\citep{leslie_2015-na}", col.names = c("Indicator", "Isla Natividad", "Maria Elena", "Punta Herrero"), format = "latex") %>% 
  kableExtra::column_spec(1, width = "10em") %>% 
  kableExtra::column_spec(2, width = "8em") %>% 
  kableExtra::column_spec(3, width = "8em") %>% 
  kableExtra::column_spec(4, width = "8em") %>% 
  kableExtra::row_spec(1 , bold = T) %>% 
  kableExtra::row_spec(6 , bold = T) %>% 
  kableExtra::row_spec(14 , bold = T) %>% 
  kableExtra::row_spec(18 , bold = T)
```

@@

```{r}
read.csv("../Data/gov_res.csv") %>% 
  knitr::kable(caption = "\\label{table:gov_res}Analysis of the fishing cooperatives based on the Social-Ecological systems framework \\citep{mcginnis_2014}.",
               col.names = c("", "Indicator", "Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  kableExtra::column_spec(1, width = "10em") %>%
  kableExtra::column_spec(2, width = "10em") %>%
  kableExtra::row_spec(1 , bold = T) %>% 
  kableExtra::row_spec(7 , bold = T) %>% 
  kableExtra::row_spec(15 , bold = T) %>% 
  kableExtra::row_spec(19 , bold = T)
```

@@

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

options(knitr.table.format = "latex")
```

@@

```{r load_pkgs}
suppressPackageStartupMessages({
  library(sp)
  library(sf)
  library(ggsn)
  library(tmap)
  library(magrittr)
  library(kableExtra)
  library(broom)
  library(cowplot)
  library(tidyverse)
})

# Load indicator functions
source(file = here::here("scripts", "abundance.R")) #Calculate abundance per transect
source(file = here::here("scripts", "richness.R")) # Calculate richness per transect
source(file = here::here("scripts", "shannon.R")) #Calculate shannon's index per transect
source(file = here::here("scripts", "biomass.R")) # Calculate fish biomass per transect
source(file = here::here("scripts", "trophic.R")) # Calculate mean trophic level per transect

# Load analysis functions
source(file = here::here("scripts", "model.R")) #fits the model
source(file = here::here("scripts", "robust_se.R")) #calculates heteroskedastic-robust Standard Errors
source(file = here::here("scripts", "plot_model.R")) #plots effect sizes
```

@@

# Load high-res mexican coastline
coastline_mx <- readRDS(here::here("data", "spatial", "coastline_mx.rds"))

@@

# Load ZRP shapefile
zrp <- st_read(dsn = here::here("data", "spatial"), layer = "ZRP", quiet = T) %>% 
  filter(Name %in% c("El Cabezo", "El Faro")) %>% 
  st_transform(crs = "+proj=longlat +datum=NAD27")

# Load Community reserve shapefile
rescom <- st_read(dsn = here::here("data", "spatial"), layer = "Res_Com", quiet = T) %>% 
  filter(Nombre == "La_Plana") %>% 
  st_transform(crs = "+proj=longlat +datum=NAD27")

@@

```{r load_data}
fish <- read.csv(here::here("data", "fish_data.csv"), strip.white = T, stringsAsFactors = F) %>% 
  filter(Comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  filter(!is.na(RC),
         RC %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
                   "Cabezo - Cabezo (Control)",
                   "Manchon - Manchon (Control)",
                   "El Faro - El Faro (Control)")) %>% 
  filter(!(RC %in% c("El Faro - El Faro (Control)") & Ano == 2012)) %>% 
  mutate(GeneroEspecie = case_when(
    GeneroEspecie == "Cephalopholis cruentatus" ~ "Cephalopholis cruentata",
    GeneroEspecie == "Haemulon macrostomun" ~ "Haemulon macrostomum",
    GeneroEspecie == "Xcochin spp" ~ "Balistes capriscus",
    GeneroEspecie == "Pterois volitans/miles" ~ "Pterois volitans",
    GeneroEspecie == "Mycteroperca spp" ~ "Mycteroperca",
    GeneroEspecie == "Chromis spp" ~ "Chromis",
    GeneroEspecie == "Damisela spp" ~ "Stegastes",
    GeneroEspecie == "Haemulon spp" ~ "Haemulon",
    GeneroEspecie == "Lancero spp" ~ "Acanthurus",
    GeneroEspecie == "Kyphosus sectatrix/incisor" ~ "Kyphosus",
    GeneroEspecie == "Sebastes spp" ~ "Sebastes",
    GeneroEspecie == "Loro spp" ~ "Scarus",
    GeneroEspecie == "Lutjanus spp" ~ "Lutjanus",
    GeneroEspecie == "Mariposa spp" ~ "Chaetodon",
    GeneroEspecie == "Pluma spp" ~ "Calamus pennatula",
    GeneroEspecie == "Sparisoma aurolineatum" ~ "Sparisoma aurofrenatum",
    GeneroEspecie == "Vieja spp" ~ "Bodianus rufus",
    TRUE ~ GeneroEspecie
  ))

invert <- read.csv(here::here("data", "invert_data.csv"), strip.white = T, stringsAsFactors = F) %>% 
  filter(Comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  filter(!is.na(RC),
         RC %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
                   "Cabezo - Cabezo (Control)",
                   "Manchon - Manchon (Control)",
                   "El Faro - El Faro (Control)")) %>% 
  filter(!(RC %in% c("El Faro - El Faro (Control)") & Ano == 2012))

load(here::here("data","conapesca_2000-2015.RData"))

cpi <- read.csv(here::here("data", "CPI.csv"), stringsAsFactors = F, strip.white = T)

@@

gob <- read.csv(here::here("data" ,"governance.csv"), strip.white = T, stringsAsFactors = T)
```

@@

```{r}
read.csv(here::here("data", "gov_ind.csv")) %>% 
  knitr::kable(caption = "\\label{table:gov_ind}Indicators used for the operationalization of the SES framework \\citep{leslie_2015-na}", col.names = c("Indicator", "Isla Natividad", "Maria Elena", "Punta Herrero"), format = "latex") %>% 
  kableExtra::column_spec(1, width = "10em") %>% 
  kableExtra::column_spec(2, width = "8em") %>% 
  kableExtra::column_spec(3, width = "8em") %>% 
  kableExtra::column_spec(4, width = "8em") %>% 
  kableExtra::row_spec(1 , bold = T) %>% 
  kableExtra::row_spec(6 , bold = T) %>% 
  kableExtra::row_spec(14 , bold = T) %>% 
  kableExtra::row_spec(18 , bold = T)
```

@@

```{r}
read.csv(here::here("data","gov_res.csv")) %>% 
  knitr::kable(caption = "\\label{table:gov_res}Analysis of the fishing cooperatives based on the Social-Ecological systems framework \\citep{mcginnis_2014}.",
               col.names = c("", "Indicator", "Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  kableExtra::column_spec(1, width = "10em") %>%
  kableExtra::column_spec(2, width = "10em") %>%
  kableExtra::row_spec(1 , bold = T) %>% 
  kableExtra::row_spec(7 , bold = T) %>% 
  kableExtra::row_spec(15 , bold = T) %>% 
  kableExtra::row_spec(19 , bold = T)
```

@@

## Summary of main findings

@@

## Limitations

## Conclusions

# Thoughts from GainesLab Meeting

## General thoughts

A nice story to tell is that reserves need to have optimal social and environmental conditions to work. But they also need to be opttimally designed, accounting for size and location. It is worth mentioning that even when these reserves did not have nice effects on the ecological or fisheries side, they were a project that brought together the community, and also gave fishers access to a network of resources facilitated by the NGO. I still need to get data that support these.

Creo que hay que agregar informacion de la variabilidad ambienta. Posiblemente SST o algo similar. Ver si se puede obtener datos de oxigeno disuelto... Tal vez algo de ENSO / ONI / NAO o algo asi

Quitar analisis de riqueza (la metodologia es restrictiva) y diversidad (aunque shannon provee une buena mezcla del cambio en abundancias y el cambio en riquezas)

revisar biomasa!

Revisar poisson o binomial negativa

Mapa de El Manchon

Las graficas y el analisis de Profit y Catches tienen que ser de langosta entera fresca solamente, y revisar que no haya habido un cambio de commodity a traves del tiempo para justificar usar solamente entera fresca... Tratar de obtener valores hasta 2017

Pedir datos biologicos hasta 2017

Distinguis profit / revenue / income

## Comentarios del Lab

Owen

Are there confounding effects because reserves are so close to control sites?

Dan

Are the data good enough to answer these questions?

Karly

Think about other social indicators that do not depend on catches or revenue from these

Alexa

Daniel

Sebas

Look at other indicators of success (adaptive capacity) in the social / governance data

Steve

Would we expect to see effects in the first five years? (ME and PH)

Are the designs good enough to see effects?

Becky

Ignacia

Perhaps interesting follow-up projects include comparing perceptions vs. reality

Estimating the benefits of being friends with an NGO

### 2018-03-20 docs/paper/Villasenor-Derbez.tex
Added manuscript to its own folder

@@

>>>>

\subsection{Summary of main findings}\label{summary-of-main-findings}

@@

\subsection{Limitations}\label{limitations}

\subsection{Conclusions}\label{conclusions}

\section{Thoughts from GainesLab
Meeting}\label{thoughts-from-gaineslab-meeting}

\subsection{General thoughts}\label{general-thoughts}

A nice story to tell is that reserves need to have optimal social and
environmental conditions to work. But they also need to be opttimally
designed, accounting for size and location. It is worth mentioning that
even when these reserves did not have nice effects on the ecological or
fisheries side, they were a project that brought together the community,
and also gave fishers access to a network of resources facilitated by
the NGO. I still need to get data that support these.

Creo que hay que agregar informacion de la variabilidad ambienta.
Posiblemente SST o algo similar. Ver si se puede obtener datos de
oxigeno disuelto\ldots{} Tal vez algo de ENSO / ONI / NAO o algo asi

Quitar analisis de riqueza (la metodologia es restrictiva) y diversidad
(aunque shannon provee une buena mezcla del cambio en abundancias y el
cambio en riquezas)

revisar biomasa!

Revisar poisson o binomial negativa

Mapa de El Manchon

Las graficas y el analisis de Profit y Catches tienen que ser de
langosta entera fresca solamente, y revisar que no haya habido un cambio
de commodity a traves del tiempo para justificar usar solamente entera
fresca\ldots{} Tratar de obtener valores hasta 2017

Pedir datos biologicos hasta 2017

Distinguis profit / revenue / income

\subsection{Comentarios del Lab}\label{comentarios-del-lab}

Owen

Are there confounding effects because reserves are so close to control
sites?

Dan

Are the data good enough to answer these questions?

Karly

Think about other social indicators that do not depend on catches or
revenue from these

Alexa

Daniel

Sebas

Look at other indicators of success (adaptive capacity) in the social /
governance data

Steve

Would we expect to see effects in the first five years? (ME and PH)

Are the designs good enough to see effects?

Becky

Ignacia

Perhaps interesting follow-up projects include comparing perceptions
vs.~reality

Estimating the benefits of being friends with an NGO
